Cycles of the logistic map

### 2013-05-02 doc/pc-ijbc.tex
doc: edited the henon map section

\newcommand{\vx}{\vct x}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\Rlam}{(R, \lambda)}

@@

A natural two-dimensional extension of the logistic map
  is the H\'enon map \cite{henon}:
\begin{equation}
  x_{k+1} = 1 + y_k - a \, {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\label{eq:henon}
\end{equation}
%
We change variable $x_k \leftarrow a x_k$, $y_k \leftarrow a y_k$, and
\begin{equation}
  x_{k+1} = a + y_k - {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\tag{$\ref{eq:henon}'$}
\label{eq:henons}
\end{equation}
%
Since neither $a$ nor $b$ is changed during the transformation,
\refeq{henon} and \refeq{henons} share the
same onset and bifurcation points
in terms of $a$ and $b$.
%
We also see that if $b \rightarrow 0$ and $a \rightarrow R$,
\refeq{henons} is reduced to the logistic map \refeq{logmaps}.

Since $y_k = b \, x_{k-1}$, we can ignore $y_k$ and work with
  cyclic polynomials of $x_k$ only,
the square free reduction is now
${x_k}^2 \rightarrow a + b x_{k-1} - x_{k+1}$.

The stability of \refeq{henons} can be found from the Jacobian matrix
\[
  \vct J_b(x_k)
  \equiv \left(
    \begin{array}{ccc}
      \partial x_{k+1}/\partial x_k & \partial x_{k+1}/\partial y_k \\
      \partial y_{k+1}/\partial x_k & \partial y_{k+1}/\partial y_k \\
    \end{array}
  \right)
  =
  \left(
    \begin{array}{ccc}
      -2 x_k & 1 \\
      b & 0
    \end{array}
  \right).
\]
The eigenvalue $\lambda$ of the composite Jacobian
$\vct J_b(x_1) \cdots \vct J_b(x_n)$ can be computed from
%
\begin{equation}
\big|\, \lambda \, \vct I - \vct J_b(x_1) \cdots \vct J_b(x_n) \,\big|
  = \lambda^2 - \Theta(\vx) \,\lambda +(-b)^n = 0,
%\tag{$\ref{eq:der}'$}
\label{eq:henonder}
\end{equation}
%
where $\Theta(\vx)$ and $(-b)^n$
are the trace and determinant of
the product matrix $\vct J_b(x_1) \cdots \vct J_b(x_n)$,
respectively \cite{hitzl}.
In a stable cycle, the magnitude of $\lambda$
cannot exceed 1; and $\lambda = +1$ or $-1$
in \refeq{henonder} corresponds to
the onset or bifurcation point, respectively.
\refeq{henonder} is the counterpart of \refeq{der}.

Since the trace of the product of two matrices is
  independent of the order,
$\Theta(\vx)$ must a cyclic polynomial of $\vx = \{ x_k \}$.
%
Thus, we can use $\Theta(\vx)$ to list \refeqs{xcp}
and then replace $\Theta(\vx)$ by $\lambda + (-b)^n/\lambda$
or $\pm\big[ 1 + (-b)^n \big]$
in \refeq{xr} to complete the solution.
%
%
%
The polynomials of $a$ and $b$
  at the onset and bifurcation points for $n$ up to 9
  are listed in the website in \refsec{end}.

@@

The odd parity, however, allows us to limit
  the set of cube-free cyclic polynomials
  to those of even degrees in $\vx$.
%
This is because \refeq{cubreplace} contains only linear and cubic terms,
and a cyclic polynomial with an odd (even) degree in $\vx$
can never be reduced to one with an even (odd) degree.
%
%Thus we only need polynomials of even degrees.
%
%We therefore have a theorem similar to \refthm{sqrfree}.
%%
%%
%\begin{theorem}
%  For the cubic map \refeq{cubic},
%  any cyclic polynomial $K(\vx)$ of
%  an $n$-cycle orbit
%  $\vct x = \{x_1, \ldots, x_n\}$
%  with an even degree in $\vct x$
%  is a linear combination of
%the even cube-free cyclic polynomials $C_p(\vx)$:
%\[
%  K(\vx) = \sum_{p \in \Pset} f_p(R) C_p(\vx),
%\]
%  where $\Pset = \{0, 11, 12, 13, \ldots, 1122, 1123, \ldots \}$ is
%  the set of indices of all even cube-free cyclic polynomials,
%  and $f_p(R)$ are polynomials of $R$.
%  \label{thm:cubfree}
%\end{theorem}
%
%
%
%
With the above change, the rest derivation is similar to that of the logistic map.
The new $\Lambda(\vx)$ should be
$\prod_{k=1}^n f'(x_k) = \prod_{k=1}^n (r - 3 {x_k}^2)$.
%
The polynomials of $r$
  at the onset and bifurcation points
%for some small $n$ are shown in \reftab{cubpolygen} (general $\lambda$)
%and \reftab{cubpoly} ($\lambda = \pm1$);
%for larger $n$ up to 8,
%we have saved the data
  are listed on the website in \refsec{end}.
%
The representative $r$ values are listed in \reftab{crval}.
%
%For complex $r$ and $\vx$,
%we have plotted \reffig{cubbulb} for regions of stability.

>>>>

\newcommand{\vx}{\vct x}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\Rlam}{(R, \lambda)}

@@

A simple two-dimensional extension of the logistic map
  is the H\'enon map \cite{henon}:
\begin{equation}
  x_{k+1} = 1 + y_k - a \, {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\label{eq:henon}
\end{equation}
%
By a change of variables
  $x_k^\mathrm{new} \equiv a \, x_k^\mathrm{old},
   y_k^\mathrm{new} \equiv a \, y_k^\mathrm{old}$,
we have, in terms of the new variables,
\begin{equation}
  x_{k+1} = a + y_k - {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\tag{$\ref{eq:henon}'$}
\label{eq:henons}
\end{equation}
%
%Since neither $a$ nor $b$ is changed during the transformation,
\refeq{henon} and \refeq{henons} share the
same onset and bifurcation points
in terms of $a$ and $b$.
%
With $b \rightarrow 0$ and $a \rightarrow R$,
\refeq{henons} is reduced to the logistic map \refeq{logmaps}.

The second equation of \refeq{henons}
allows us to replace $y_k$ by $b \, x_{k-1}$,
  such that we can work with the cyclic polynomials of $x_k$ only.
The square-free reduction is now
  ${x_k}^2 \rightarrow a + b x_{k-1} - x_{k+1}$.

The counterpart of \refeq{der}
  can be found from the Jacobian of $n$th iterate of \refeq{henons}
\[
  \vct J^n
  \equiv
  \left(
    \begin{array}{ccc}
      \partial x_{n+1}/\partial x_1 & \partial x_{n+1}/\partial y_1 \\
      \partial y_{n+1}/\partial x_1 & \partial y_{n+1}/\partial y_1 \\
    \end{array}
  \right),
\]
and the magnitude of both eigenvalues of the matrix must be less than 1.
By the chain rule, we have
  $\vct J^n (x_1) = \vct J^1(x_1) \cdots \vct J^1(x_n)$,
  where
$\vct J^1(x_k)
  =
  \left(
    \begin{array}{cc}
      -2 x_k & 1 \\
      b & 0
    \end{array}
  \right)$.
Thus, the eigenvalues $\lambda$ of $\vct J^n$ can be found from
%
\begin{equation}
\big|\, \lambda \, \vct I - \vct J^n(x_1) \,\big|
  = \lambda^2 - \Theta(\vx) \,\lambda +(-b)^n = 0,
%\tag{$\ref{eq:der}'$}
\label{eq:henonder}
\end{equation}
%
where $\Theta(\vx)$ and $(-b)^n$
  are the trace and determinant of $\vct J^n(x_1)$,
  respectively \cite{hitzl}.
The onset and bifurcation points
  correspond to $\lambda = +1$ and $-1$, respectively

Using the matrix identity $\Tr( \vct A \vct B) = \Tr(\vct B \vct A)$,
  we have
    $\Tr \, \vct J^n(x_1) = \cdots = \Tr \, \vct J^n(x_n)$,
i.e., $\Theta(\vx)$ is a cyclic polynomial of $\vx = \{ x_k \}$.
%e.g., %$\Theta(\vx) = 4 x_1 x_2 + 2 \, b$ for $n = 2$ and
%$\Theta(\vx) = -8 x_1 x_2 x_3 - 2 \, b (x_1+ x_2+ x_3)$ for $n = 3$.
%
Thus, we can expand $\Theta(\vx) \, C_p(\vx)$
  in terms of square-free cyclic polynomials,
  and then replace $\Theta(\vx)$ by $\lambda + (-b)^n/\lambda$
  in the final characteristic equation to obtain the desired polynomials
  (the procedure is similar to that in \refsec{algo} for the logistic map).
%
%
%
The polynomials of $a$ and $b$
  at the onset and bifurcation points for $n$ up to 9
  are listed in the website in \refsec{end}.

@@

The odd parity, however, allows us to limit
  the set of cube-free cyclic polynomials
  to those of even degrees in $\vx$.
%
This is because \refeq{cubreplace} contains only linear and cubic terms,
and a cyclic polynomial with an odd (even) degree in $\vx$
can never be reduced to one with an even (odd) degree.
%
%Thus, we only need polynomials of even degrees.
%
%We therefore have a theorem similar to \refthm{sqrfree}.
%%
%%
%\begin{theorem}
%  For the cubic map \refeq{cubic},
%  any cyclic polynomial $K(\vx)$ of
%  an $n$-cycle orbit
%  $\vct x = \{x_1, \ldots, x_n\}$
%  with an even degree in $\vct x$
%  is a linear combination of
%the even cube-free cyclic polynomials $C_p(\vx)$:
%\[
%  K(\vx) = \sum_{p \in \Pset} f_p(R) C_p(\vx),
%\]
%  where $\Pset = \{0, 11, 12, 13, \ldots, 1122, 1123, \ldots \}$ is
%  the set of indices of all even cube-free cyclic polynomials,
%  and $f_p(R)$ are polynomials of $R$.
%  \label{thm:cubfree}
%\end{theorem}
%
%
%
%
With the above change, the rest derivation is similar to that of the logistic map.
The new $\Lambda(\vx)$ should be
$\prod_{k=1}^n f'(x_k) = \prod_{k=1}^n (r - 3 {x_k}^2)$.
%
The polynomials of $r$
  at the onset and bifurcation points
%for some small $n$ are shown in \reftab{cubpolygen} (general $\lambda$)
%and \reftab{cubpoly} ($\lambda = \pm1$);
%for larger $n$ up to 8,
%we have saved the data
  are listed on the website in \refsec{end}.
%
The representative $r$ values are listed in \reftab{crval}.
%
%For complex $r$ and $\vx$,
%we have plotted \reffig{cubbulb} for regions of stability.

### 2013-05-02 doc/pc.tex
doc: edited the henon map section

We now extend the method to the H\'enon map \cite{henon}:
\begin{equation}
  x_{k+1} = 1 + y_k - a \, {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\label{eq:henon}
\end{equation}
%
We change variable $x_k \leftarrow a x_k$, $y_k \leftarrow a y_k$, and
\begin{equation}
  x_{k+1} = a + y_k - {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\tag{$\ref{eq:henon}'$}
\label{eq:henons}
\end{equation}
%
Since neither $a$ nor $b$ is changed during the transformation,
\refeq{henon} and \refeq{henons} share the
same onset and bifurcation points
in terms of $a$ and $b$.
%
We also see that if $b \rightarrow 0$ and $a \rightarrow R$,
\refeq{henons} is reduced to the logistic map \refeq{logmaps}.

Since $y_k = b \, x_{k-1}$, we can ignore $y_k$ and work with
  cyclic polynomials of $x_k$ only,
the square free reduction is now
${x_k}^2 \rightarrow a + b x_{k-1} - x_{k+1}$.

The stability of \refeq{henons} can be found from the Jacobian matrix
\[
  \vct J_b(x_k)
  \equiv \left(
    \begin{array}{ccc}
      \partial x_{k+1}/\partial x_k & \partial x_{k+1}/\partial y_k \\
      \partial y_{k+1}/\partial x_k & \partial y_{k+1}/\partial y_k \\
    \end{array}
  \right)
  =
  \left(
    \begin{array}{ccc}
      -2 x_k & 1 \\
      b & 0
    \end{array}
  \right).
\]
The eigenvalue $\lambda$ of the composite Jacobian
$\vct J_b(x_1) \cdots \vct J_b(x_n)$ can be computed from
%
\begin{equation}
\big|\, \lambda \, \vct I - \vct J_b(x_1) \cdots \vct J_b(x_n) \,\big|
  = \lambda^2 - \Theta(\vx) \,\lambda +(-b)^n = 0,
%\tag{$\ref{eq:der}'$}
\label{eq:henonder}
\end{equation}
%
where $\Theta(\vx)$ and $(-b)^n$
are the trace and determinant of
the product matrix $J_b(x_1) \cdots J_b(x_n)$,
respectively \cite{hitzl}.
In a stable cycle, the magnitude of $\lambda$
cannot exceed 1; so we replace $\lambda$ by $+1$ or $-1$
in \refeq{henonder}
to obtain the onset or bifurcation point, respectively.
\refeq{henonder} is the counterpart of \refeq{der}.

Since cyclically rotating matrices in a product does not alter the trace,
$\Theta(\vx)$ is a cyclic polynomial of $\vx = \{ x_k \}$.
%e.g.,
%$\Theta(\vx) = 4 x_1 x_2 + 2 \, b$ for $n = 2$ and
%$\Theta(\vx) = -8 x_1 x_2 x_3 - 2 \, b (x_1+ x_2+ x_3)$ for $n = 3$.
%
Thus, we can use $\Theta(\vx)$ to list \refeqs{xcp}
and then replace $\Theta(\vx)$ by $\lambda + (-b)^n/\lambda$
or $\pm\big[ 1 + (-b)^n \big]$
in \refeq{xr} to complete the solution.

>>>>

A simple two-dimensional extension of the logistic map
  is the H\'enon map \cite{henon}:
\begin{equation}
  x_{k+1} = 1 + y_k - a \, {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\label{eq:henon}
\end{equation}
%
By a change of variables
  $x_k^\mathrm{new} \equiv a \, x_k^\mathrm{old},
   y_k^\mathrm{new} \equiv a \, y_k^\mathrm{old}$,
we have, in terms of the new variables,
\begin{equation}
  x_{k+1} = a + y_k - {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\tag{$\ref{eq:henon}'$}
\label{eq:henons}
\end{equation}
%
%Since neither $a$ nor $b$ is changed during the transformation,
\refeq{henon} and \refeq{henons} share the
same onset and bifurcation points
in terms of $a$ and $b$.
%
With $b \rightarrow 0$ and $a \rightarrow R$,
\refeq{henons} is reduced to the logistic map \refeq{logmaps}.

The second equation of \refeq{henons}
allows us to replace $y_k$ by $b \, x_{k-1}$,
  such that we can work with the cyclic polynomials of $x_k$ only.
The square-free reduction is now
  ${x_k}^2 \rightarrow a + b x_{k-1} - x_{k+1}$.

The counterpart of \refeq{der}
  can be found from the Jacobian of $n$th iterate of \refeq{henons}
\[
  \vct J^n
  \equiv
  \left(
    \begin{array}{ccc}
      \partial x_{n+1}/\partial x_1 & \partial x_{n+1}/\partial y_1 \\
      \partial y_{n+1}/\partial x_1 & \partial y_{n+1}/\partial y_1 \\
    \end{array}
  \right),
\]
and the magnitude of both eigenvalues of the matrix must be less than 1.
By the chain rule, we have
  $\vct J^n (x_1) = \vct J^1(x_1) \cdots \vct J^1(x_n)$,
  where
$\vct J^1(x_k)
  =
  \left(
    \begin{array}{cc}
      -2 x_k & 1 \\
      b & 0
    \end{array}
  \right)$.
Thus, the eigenvalues $\lambda$ of $\vct J^n$ can be found from
%
\begin{equation}
\big|\, \lambda \, \vct I - \vct J^n(x_1) \,\big|
  = \lambda^2 - \Theta(\vx) \,\lambda +(-b)^n = 0,
%\tag{$\ref{eq:der}'$}
\label{eq:henonder}
\end{equation}
%
where $\Theta(\vx)$ and $(-b)^n$
  are the trace and determinant of $\vct J^n(x_1)$,
  respectively \cite{hitzl}.
The onset and bifurcation points
  correspond to $\lambda = +1$ and $-1$, respectively

Using the matrix identity $\Tr( \vct A \vct B) = \Tr(\vct B \vct A)$,
  we have
    $\Tr \, \vct J^n(x_1) = \cdots = \Tr \, \vct J^n(x_n)$,
i.e., $\Theta(\vx)$ is a cyclic polynomial of $\vx = \{ x_k \}$.
%e.g., %$\Theta(\vx) = 4 x_1 x_2 + 2 \, b$ for $n = 2$ and
%$\Theta(\vx) = -8 x_1 x_2 x_3 - 2 \, b (x_1+ x_2+ x_3)$ for $n = 3$.
%
Thus, we can expand $\Theta(\vx) \, C_p(\vx)$
  in terms of square-free cyclic polynomials,
  and then replace $\Theta(\vx)$ by $\lambda + (-b)^n/\lambda$
  in the final characteristic equation to obtain the desired polynomials
  (the procedure is similar to that in \refsec{algo} for the logistic map).
%
%
%
%The polynomials of $a$ and $b$
%  at the onset and bifurcation points for $n$ up to 9
%  are listed in the website in \refsec{end}.
